### 2017-05-29 paper/invariants-martingales.tex
edited prelims

%\vspace{-1em}
\section{Invariants and Ranking Supermartingales}\label{sec:invm}
%\vspace{-0.5em}

In this section we recall known methods and constructs for solving the 
qualitative termination and reachability questions for \APP{s}, namely  
linear invariants and ranking supermartingales. 
We also demonstrate that these methods are not sufficient to address the 
quantitative variants of these questions (i.e., probabilistic termination).
In order to discuss the necessary concepts, we recall the basics of 
martingales, which is relevant for both this and subsequent sections. 
%Since the definitions of this section are from the literature we relegate
%relevant illustrations through example to the Appendix.

%While the core results of this section were already proven in the previous 
%literature, the version we sometimes present slight generalizations of these 
%result, which make them applicable to a larger class of programs. We discuss 
%the nature of these generalizations where appropriate.

%\vspace{-1em}
\subsection{Pure Invariants}
%\vspace{-0.5em}

Invariants are a vital element of many program analysis techniques. 
Intuitively, invariants are maps assigning to each 
program location $\loc$ of some pCFG a predicate which is guaranteed to hold 
whenever $\loc$ is 
entered. To avoid confusion with stochastic invariants, 
that we introduce later, we call these standard invariants \emph{pure invariants}.

\smallskip
\begin{definition}[Linear Predicate Map (LPM) and Pure Invariant] We define the 
following:
\begin{compactenum}
\item
A \emph{linear predicate map (LPM)} for an \APP{} $P$ is a function $I$ 
assigning to each location $\loc$ of the pCFG $\pCFG_P$ a propositionally 
linear predicate $I(\loc)$ over the set of program variables of $P$.
\item
A \emph{pure linear invariant} (or just a pure invariant) for an \APP{} $P$ is 
a linear predicate map $\inv$ for $P$ with
the following property: for each location $\loc$ of $\pCFG_P$ and each finite 
path $(\loc_0,\vec{x}_0),\cdots,(\loc_n,\vec{x}_n)$ such that 
$(\loc_0,\vec{x}_0)=(\locinit,\vecinit)$ and $\loc_n = \loc$ it holds
$\vec{x}_n\models \inv(\loc)$.
\end{compactenum}
%%A 
%%probabilistic LPM for $P$ is a tuple $(\pinv,\pinfunc)$ such that $\pinv$ is 
%%a 
%%pure LPM for $P$ and $\pinfunc$ is a function assigning to each location 
%%$\loc$ 
%%a probability $\pi(\loc)\in[0,1]$.
\end{definition}

%A special class of LPMs are program invariants, i.e. maps assigning to each 
%program location $\loc$ an PLP which is guaranteed to hold whenever $\loc$ is 
%entered.
%Invariants are a standard construct used in automated program analysis. 
%To avoid confusion with probabilistic invariants, we call the 
%standard invariants \emph{pure}.
%
%\begin{definition}[Pure Linear Invariant]
%A \emph{pure linear invariant} (or just a pure invariant) for an \APP{} $P$ is 
%%a function map $\inv$  assigning to each location $\loc$ of the pCFG 
%%$\pCFG_P$ 
%%a propositionally 
%%linear predicate $I(\loc)$ over the set of program variables of $P$ 
%a linear predicate map $\inv$ for $P$ with
%the following property: for each location $\loc$ of $\pCFG_P$ and each finite 
%path $(\loc_0,\vec{x}_0),\cdots,(\loc_n,\vec{x}_n)$ such that 
%$(\loc_0,\vec{x}_0)=(\locinit,\vecinit)$ and $\loc_n = \loc$ it holds
%$\val_{\vec{x}_i}\models \inv(\loc)$.
%\end{definition}

%\vspace{-1em}
\subsection{Supermartingales}
%\vspace{-0.5em}

%Submartingales and their counterparts, supermartingales, are a standard tool 
%of 
(Super)martingales, are a standard tool of 
probability theory apt for analyzing probabilistic objects arising in computer 
science, from automata-based models~\cite{BKKNK:pMC-zero-reachability} to 
general probabilistic 
programs~\cite{SriramCAV,HolgerPOPL,CFNH16:prob-termination,CFG16:positivstellensatz-arxiv,BEFH16:doob-decomposition}.

%\PN{Maybe move this sentence to intro.} 
Let us first recall basic
definitions and results related to supermartingales, which we need in our 
analysis.

\smallskip\noindent{\bf Conditional Expectation.} 
%The notion of conditional expectation 
%tightly connected to martingales. We present only a brief informal exposition 
%here, for formal treatment see, e.g.~\cite[Chapter 9]{Williams:book}.
Let $(\Omega,\mathcal{F},\probm)$ be a probability space, 
$X\colon\Omega\rightarrow 
\Rset$ an $\mathcal{F}$-measurable function, and $\mathcal{F}'\subseteq 
\mathcal{F}$ sub-sigma-algebra of $\mathcal{F}$. The \emph{conditional 
expectation} of $X$ given $\mathcal{F}'$ is an $\mathcal{F}'$-measurable random 
variable denoted by $\E[X| \mathcal{F}']$ which satisfies, for each set $A\in 
\mathcal{F}'$, the following: 
\begin{equation}
\label{eq:cond-exp}
\E[X\cdot 1_A] = \E[\E[X|\mathcal{F}]\cdot 1_A],
\end{equation}
where $1_A \colon \Omega\rightarrow \{0,1\}$ is an \emph{indicator function} of 
$A$, i.e. function returning $1$ for 
each $\omega\in A$ and $0$ for each $\omega\in \Omega\setminus A$. Note that 
the left hand-side of~\eqref{eq:cond-exp} intuitively represents the expected 
value 
of $X(\omega)$ with domain restricted to $A$.

Recall that in context of probabilistic programs we work with probability 
spaces of the form $(\Omega,\natfilt,\probm^\sigma)$, where $\Omega$ is a 
set of runs in some $\pCFG$ and $\mathcal{F}$ is (the smallest) sigma-algebra 
such that all the functions $\cfg{\sigma}{i}$, where $i\in \Nset_0$ and 
$\sigma$ is a scheduler, are $\natfilt$-measurable. In such a setting we can 
also consider sub-sigma-algebras $\natfilt_i$, $i\in \Nset_0$, of 
$\natfilt$, where $\natfilt_i$ is the smallest sub-sigma-algebra of 
$\natfilt$ such that all the functions $\cfg{\sigma}{j}$, $0\leq j \leq 
i$, are $\natfilt_i$-measurable. Intuitively, each set $A$ belonging to such 
an $\natfilt_i$ consists of runs whose first $i$ steps satisfy some 
property, and the probability space $(\Omega,\natfilt_i,\probm^\sigma)$ 
allows us to reason about probabilities of certain events happening in the 
first 
$i$ steps of program execution. 
%(note that since $\natfilt_i \subseteq 
%\natfilt$, we can view $\probm^\sigma$ also as a function of type 
%$\natfilt_i\rightarrow [0,1]$). 
Then, for each $A\in \natfilt_i$, the 
value $\E[\E[X|\natfilt_i]\cdot 1_A]$ represents the expected value of 
$X(\run)$ for the randomly generated run $\run$ provided that we restrict to 
runs whose
prefix of length $i$ satisfies the property given by $A$. 
Note that the sequence $\natfilt_0,\natfilt_1,\natfilt_2,\dots$ forms a 
filtration of $\natfilt$, which we call a \emph{canonical filtration}.

%%\paragraph*{Basics of Supermartingales}

\smallskip
\begin{definition}[Supermartingale]
Let $(\Omega,\mathcal{F},\probm)$ be a probability space and 
$\{\mathcal{F}_i\}_{i=0}^{\infty}$ a filtration of $\mathcal{F}$. A sequence of 
random variables $\{X_i\}_{i=0}^{\infty}$ is a \emph{supermartingale} w.r.t. 
filtration $\{\mathcal{F}_i\}_{i=0}^{\infty}$ if it satisfies these conditions:
\begin{compactenum}
\item  The process $\{X_i\}_{i=0}^{\infty}$ is \emph{adapted} to 
$\{\mathcal{F}_i\}_{i=0}^{\infty}$, i.e. for all $i\in \Nset_0$ it holds that 
$X_i$ is $\mathcal{F}_i$-measurable.
\item For all $i\in \Nset_0$ it holds $\E[|X_i|]<\infty$.
\item For all $i\in \Nset_0$ it holds 
\begin{equation}
\label{eq:supermart-def}
\E[X_{i+1}|\mathcal{F}_i] \leq X_i.
\end{equation}

A supermartingale $\{X_i\}_{i=0}^{\infty}$ has $c$-bounded differences, where 
$c\geq 0$, if $|X_{i+1}-X_i|<c$ for all $i\in \Nset_0$
\end{compactenum}
%A sequence of 
%random variables $\{X_i\}_{i=0}^{\infty}$ is a \emph{submartingale} w.r.t.  
%$\{\mathcal{F}_i\}_{i=0}^{\infty}$ if the process $\{-X_i\}_{i=0}^{\infty}$ is 
%a supermartingale w.r.t. the same filtration.
\end{definition}

Intuitively, a supermartingale is a stochastic process whose average value is 
guaranteed not to rise as time evolves, even if some information on the past 
evolution of the process is revealed. 
%Symmetrically, a submartingale is a 
%process whose value does not drop on average. 
We often need to work with supermartingales whose value 
is guaranteed to decrease on average, until a certain condition is 
satisfied. The point in time in which such a condition is satisfied is called a 
\emph{stopping time}.

\smallskip
\begin{definition}[Stopping time]
Let $(\Omega,\mathcal{F},\probm)$ be a probability space and 
$\{\mathcal{F}_i\}_{i=0}^{\infty}$ a filtration. A random variable $\stime 
\colon 
\Omega\rightarrow \Nset_0$ is called a \emph{stopping time} w.r.t. 
$\{\mathcal{F}_i\}_{i=0}^{\infty}$ if %$\stime$ is adapted to this filtration 
%and if 
for all $j\in \Nset_0$ the set $\{\omega\in \Omega\mid \stime(\omega)\leq j\}$ 
belongs to $\mathcal{F}_j$.
\end{definition}

In particular, for each set of configurations $C$ the reachability time 
$\treach{C}$ of $C$ is a stopping time w.r.t. the canonical filtration, since 
at each time $j$ we can decide whether $\treach{C}>j$ or not by looking at the 
prefix of a run of length $j$. 
%When $\Omega$ is a set of runs of some $\pCFG_P$, we can think of 
%$\stime$ as representing a 
%certain condition which can or cannot hold in a given time instance of the 
%execution of $P$, where $\stime(\run)$ represents the first point in time 
%during the execution represented by $\run$ in which $P$ holds. The condition 
%$\{\omega\in \Omega\mid \stime(\omega)\leq j\}\in \mathcal{F}_j$ then states 
%that whether $P$ holds at step $j$ of a run can be determined by looking at 
%the 
%prefix of the run of length $j$. In other words, the stopping condition cannot 
%``talk'' about future events. The typical stopping time we use in the context 
%of \APP{s} is the \emph{reachability} time $\treach{C}$ for some set of 
%configurations $C$, where $\treach{C}(\run)$ is the first point in time when 
%the current configuration on $\run$ is in $C$. In particular, if $C$ is the 
%set 
%of all configurations $(\loc,\vec{x})$ such that $\loc=\loc^\lout_P$ (the 
%terminal location of $\pCFG_P$), then we denote $\treach{C}$ by $\ttime$ and 
%call it a \emph{termination time} of $P$.
%We combine stopping conditions with martingales to get 
%\emph{$\eps$-decreasing} 
%supermartingales. %
Finally, we recall the fundamental notion of a ranking 
supermartingale.

\smallskip
\begin{definition}[Ranking supermartingale]
\label{def:ranking}
Let $(\Omega,\mathcal{F},\probm)$ be a probability space, 
$\{\mathcal{F}_i\}_{i=0}^{\infty}$ a filtration of $\mathcal{F}$, $\stime$ a 
stopping time w.r.t. that filtration, and $\eps\geq 
0$. 
A supermartingale $\{X_i\}_{i=0}^{\infty}$ (w.r.t. 
$\{\mathcal{F}_i\}_{i=0}^{\infty}$) is \emph{$\eps$-decreasing} until 
$\stime$ 
if it 
satisfies 
the following additional condition: for all $i\in \Nset_0$ it holds
\begin{equation}
\label{eq:ranking-sup}
\E[X_{i+1}|\mathcal{F}_i] \leq X_i - \eps\cdot\vec{1}_{\stime > i}.
\end{equation}

Further, $\{X_i\}_{i=0}^{\infty}$ is an \emph{$\eps$-ranking} supermartingale 
($\eps$-RSM) for $\stime$ if it 
is $\eps$-decreasing until $\stime$ and for 
each $\omega\in\Omega, j\in\Nset_0$ it holds $\stime(\omega)>j \Rightarrow 
X_j(\omega)\geq 0$.
%is a supermartingale w.r.t. $\{\mathcal{F}_i\}_{i=0}^{\infty}$. 
%A supermartingale $\{X_i\}_{i=0}^{\infty}$ is 
%is $\eps$-repulsing if for all $i\in \Nset_0$ it holds
%\begin{equation}
%\label{eq:repulsing-sub}
%\E[X_{i+1}|\mathcal{F}_i] \leq X_i - \eps\cdot\vec{1}_{\stime > i}.
%\end{equation}
\end{definition}
Intuitively, if $\stime$ is the reachability time $\treach{C}$ of some set $C$, 
then the previous definition requires that an $\eps$-ranking supermartingale 
must decrease by at least $\eps$ on average up to the point when $C$ is 
reached for a first time. After that, it must not increase (on average).
The above definition is a bit more general than the standard one in the
literature as we also consider reachability as opposed to only termination.

%Intuitively, until the condition specified by $\stime$ is satisfied, ranking 
%supermartingale is non-negative and at the same time decreasing by at least 
%$\eps$ on average, forming a probabilistic analogue of the classical ranking 
%functions.

%We finish this part by noting that for some applications it is preferable to 
%work with supermartingales that have \emph{bounded differences.}

%\begin{definition}
%A supermartingale $\{X_i\}_{i=0}^{\infty}$ has $c$-bounded differences, where 
%$c\geq 0$, if $|X_{i+1}-X_i|<c$ for all $i\in \Nset_0$.
%\end{definition}
%Note that $\{X_i\}_{i=0}^{\infty}$ being an $\eps$-repulsing supermartingale 
%for 
%$\stime$ is 
%\emph{not} equivalent to $\{-X_i\}_{i=0}^{\infty}$ being an $\eps$-ranking 
%supermartingale for $\stime$.

%We typically work with $\eps$-ranking supermartingales whose initial value is 
%positive and $\eps$-repulsing supermartingales whose initial value is 
%negative. 
%Then, intuitively an $\eps$-ranking supermartingale is driven towards negative 
%values by at least $\eps$ in each step (on average) until it indeed becomes 
%negative, while an $\eps$-repulsing submartingale is driven \emph{away} from 
%positive values by at least $\eps$ per step, as long as it stays non-positive 
%(we cannot rule out that it eventually becomes positive, as the decrease by 
%$\eps$ is guaranteed only on average). 
%%Note that $\{X_i\}_{i=0}^{\infty}$ being an $\eps$-repulsing submartingale is 
%%\emph{not} equivalent to $\{-X_i\}_{i=0}^{\infty}$ being an $\eps$-ranking 
%%supermartingale.

\smallskip\noindent{\bf Martingales in Program Analysis.}
In the context of \APP{} analysis, we consider a special type of 
supermartingales given as functions of the current values of program variables. 
In this paper we focus on the case when these functions are \emph{linear}.

\smallskip
\begin{definition}[Linear Expression Map]
A \emph{linear expression map (LEM)} for an \APP{} $P$ is a function $\lem$ 
assigning to each program location $\loc$ of $\pCFG_P$ an affine expression 
$\lem(\ell)$ over the program variables of $P$.
\end{definition}

Each LEM $\lem$ and location $\loc$ determines an affine function $\lem(\loc)$ 
which takes as an argument an $n$-dimensional vector, where $n$ is the number 
of distinct variables in $P$. We use $\lem(\loc,\vec{x})$ as a shorthand 
notation for $\lem(\loc)(\vec{x})$. 
Martingales for \APP{} analysis are defined via a standard notion of 
pre-expectation~\cite{SriramCAV}. Intuitively, a pre-expectation of $\lem$ is a 
function which for each configuration $(\loc,\vec{x})$ returns the maximal
expected value of $\lem$ after one step is made from this configuration, where 
the maximum is taken over all possible non-deterministic choices.

\smallskip
\begin{definition}[Pre-Expectation]
Let $P$ be an \APP{} such that $\pCFG_P = 
(\locs,\pvars,\locinit,\vecinit,\transitions,\probdist,\guards)$ and let $\lem$ 
a 
linear expression map 
for~$P$. 
%Let $\locs$ 
%be the set of locations of $\pCFG_P$ and $\pvars$ its set of program 
%variables. 
The 
pre-expectation of $\lem$ is a function $\preexp{\lem}\colon \locs\times 
\Rset^{|\pvars|} \rightarrow \Rset$ defined as follows:
\begin{compactitem} %\itemsep1pt \parskip0pt \parsep0pt
\item 
if $\loc$ is a probabilistic location, then
$$\preexp{\lem}(\loc,\vec{x}):=\sum_{(\loc,1,\id_1,\loc')\in\transitions} 
Pr_{\loc}\left((\loc,1,\id_1,\loc')\right)\cdot
 \lem(\loc',\vec{x});$$
\item 
 if $\loc$ is a non-deterministic location, then
$$\preexp{\lem}(\loc,\vec{x}):=
\max_{(\loc,1,\id_1,\loc')\in\transitions}\lem(\loc',\vec{x});$$

%\item 
%$\mathrm{pre}_\eta(\loc,\mathbf{x}):=\min_{(\loc,id,\ell')\in\transitions}\eta\left(\loc',\mathbf{x}\right)$
% if $\loc$ is an angelic location;
\item 
if $\loc$ is a deterministic location, then for each $\vec{x}$ the value 
$\preexp{\lem}(\loc,\vec{x})$ is determined as follows: there is exactly one 
transition
$\tau=(\loc,j,\up,\loc')$ such that $\vec{x}\models G(\tau)$. We distinguish 
three cases:
\begin{compactitem}
\item If $\up\colon \Rset^{|\pvars|}\rightarrow \Rset$ is a function, then 
$$\preexp{\lem}(\loc,\vec{x}):= \lem(\loc',\vec{x}(j\leftarrow \up(\vec{x}))).$$
\item If $\up$ is a distribution $d$, then $$ \preexp{\lem}(\loc,\vec{x}):= 
\lem(\loc',\vec{x}(j\leftarrow \E[d])),$$ where $\E[d]$ is the expected value of the 
distribution $d$.
\item 
If $\up$ is a set, then $$ \preexp{\lem}(\loc,\vec{x}):= \max_{a\in\up}
\lem(\loc',\vec{x}(j \leftarrow a)).$$
\end{compactitem}
%$\mathrm{pre}_\eta(\loc,\mathbf{x}):=\eta(\loc',\expv_{\rvars}\left(f(\vec{x},\mathbf{r})\right))$
% if $\loc$ is a deterministic location, $(\loc,f,\loc')\in\transitions$ and 
%$\mathbf{x}\in G(\loc,f,\loc')$, where 
%$\expv_{\rvars}\left(f(\vec{x},\mathbf{r})\right)$ is the expected value of 
%$f(\vec{x},\cdot)$. %on $\rvars$ if $\vec{x}$ is treated as a constant vector.
\end{compactitem}
%\end{definition}
\end{definition}

\smallskip
\begin{definition}(Linear Ranking Supermartingale)
\label{def:lrsm}
Let $P$ be an \APP{} such that $\pCFG_P = 
(\locs,\pvars,\locinit,\vecinit,\transitions,\probdist,\guards)$, let $\inv$ be 
a linear predicate map and let 
$\confset\subseteq \locs \times \Rset^{|\pvars|}$ be some set of 
configurations. 
%\begin{compactenum}
%\item
A linear $\eps$-ranking supermartingale ($\eps$-LRSM) for $\confset$ 
supported by $\inv$ is a 
linear expression map $\lem$ for $P$ such that for all configurations 
$(\loc,\vec{x})$ of $\pCFG_P$ with
$(\loc,\vec{x})\not\in \confset$ and $\vec{x}\models \inv(\loc)$ the following 
two conditions hold:
\begin{compactitem}
%\item For all $(\loc,\vec{x})\in \locs\times\Rset^{|\pvars|}$ it holds
%\item For all  it holds 
\item
$\lem(\loc,\vec{x})\geq 0$
\item 
$\preexp{\lem}(\loc,\vec{x}) \leq \lem(\loc,\vec{x})-\eps$ 
\end{compactitem}
%\item
%\end{compactenum}
A linear $\eps$-ranking supermartingale supported by $\inv$ has 
$c$-bounded differences if for each $(\loc,\vec{x})$ such that $\vec{x}\models 
\inv(\loc)$ and each configuration $(\loc',\vec{x}')$ such that $(\loc,\vec{x}) 
(\loc',\vec{x}')$ is a path in $\pCFG_P$ it holds 
$|\lem(\loc,\vec{x})-\lem(\loc',\vec{x}')|\leq c$. 
\end{definition}

The relationship between $\eps$-LRSM in \APP{s}, (pure) invariants, and almost-sure termination 
is summarized in the following theorem. 
\smallskip
\begin{theorem}[{\cite[Theorem 1]{CFNH16:prob-termination}}]
\label{thm:old-ranking}
Let $P$ be an \APP{}, $\sigma$ a scheduler, and 
$(\Omega,\natfilt,\probm^\sigma)$ the corresponding probability space. Further, 
let $C$ be the set of terminating configurations of $\pCFG_P$ (i.e., the termination
location is reached), such that there exist an $\eps>0$ and an $\eps$-linear ranking 
supermartingale $\lem$ supported by a pure invariant $I$. 
%%for $\treach{C}$. 
Then 
\begin{compactenum}
\item
$\probm^{\sigma}(\ttime<\infty)=1$, i.e. termination is ensured almost-surely.
\item
$\E^{\sigma}[\ttime]<\lem(\locinit,\vecinit)/\eps$.
%\item
%If $\{X_i\}_{i=0}^{\infty}$  has $c$-bounded differences for some $c$, then 
%there exists a concentration bound $B$ for $\treach{C}$ such that $B\leq 
%({E^{\sigma}[X_0]}/{\eps})+1$.
\end{compactenum}
\end{theorem}

The previous result shows that if there exists  an $\eps$-LRSM 
supported by a pure invariant $I$, for $\eps>0$, then under each scheduler 
termination is ensured almost-surely.
We now demonstrate that pure invariants, though effective for almost-sure 
termination, are ineffective to answer probabilistic termination questions.

%\begin{figure}[t]
\lstset{language=affprob}
\lstset{tabsize=3}
\newsavebox{\nonterm}
\begin{lrbox}{\nonterm}
\begin{lstlisting}[mathescape]
$x:=30,y:=20$
while $y\geq 0$ do
	$x:=x+$sample$(\mathrm{Uniform}[-\frac{1}{4},1])$
	$y:=y+$sample$(\mathrm{Uniform}[-1,\frac{1}{4}])$
	while $x \leq 0$ do skip od
od
\end{lstlisting}
\end{lrbox}
\begin{figure}[t]
%%\centering
\usebox{\nonterm}
\begin{tikzpicture}[x = 1.7cm]
%\node[det] (det1) at (0,0) {$\loc_0$};
%\node[det] (while) at (1.5,-1.5)  {$\loc_1$};
%\node[det] (fin) at (0,-3) {$\loc_5$};
%\node[ran] (prob) at (3,-1.5) {$\loc_2$};
%\node[ang] (angel) at (3,0) {$\loc_3$};
%\node[dem] (demon) at (3,-3)  {$\loc_4$};
%
%\draw[tran] (det1) to node[auto, swap] {x:=0} (while);
%\draw[tran] (while) to node[font=\scriptsize,draw, fill=white, 
%rectangle,pos=0.3] {$x<0$} (fin);
%\draw[tran, loop, looseness = 5, in =-65, out = -115] (fin) to (fin);
%\draw[tran] (while) to node[font=\scriptsize,draw, fill=white, 
%rectangle,pos=0.3] {$x\geq 0$} (prob);
%\draw[tran] (prob) to node[font=\scriptsize,draw, fill=white, 
%rectangle,pos=0.3, inner sep = 1pt] {$\frac{6}{10}$} (angel);
%\draw[tran] (prob) to node[font=\scriptsize,draw, fill=white, 
%rectangle,pos=0.3, inner sep = 1pt] {$\frac{4}{10}$} (demon);
%\draw[tran] (demon) -- node[auto] {x:=x+1} (demon-|while) -- (while);
%\draw[tran] (angel) -- node[auto, swap] {x:=x+1} (angel-|while) -- (while);
%\draw[tran] (demon) to node[auto] {x:=x-1} (while);
%\draw[tran] (angel) to node[auto, swap] {x:=x-1} (while);

\node[det] (while) at (1.5,0)  {$\loc_0$};
%%\node[det] (fin) at (0,-3) {};
\node[det] (p1) at (2.5,0) {$\loc_1$};
\node[det] (p2) at (3.5,0) {$\loc_2$};

\node[det] (lp) at (4.5,0) {$\loc_3$};
%\node[ang] (angel) at (3,0) {$\loc_3$};
%\node[dem] at (3,-3) (demon) {$\loc_4$};
%\node[det] (aif) at (2,-0.075) {$\loc_5$};
%\node[det] (aelse) at (3,-0.8) {$\loc_6$};
%\node[det] (dif) at (2,-3) {$\loc_7$};
%\node[det] (delse) at (3,-2.2) {$\loc_8$};
%\node[det] (det1) at (0,0) {$\loc_0$};
%\node[det] (while) at (1.5,-1.5)  {};
\node[det] (fin) at (0,0) {$\loc_4$};
%\node[ran] (prob) at (3,-1.5) {};
%\node[ang] (angel) at (3,0) {};
%\node[dem] at (3,-3) (demon) {};
%
%%\draw[tran] (det1) to node[auto, swap] {x:=0} (while);
\node (dum1) at (0,0.8) {};
\node (dum2) at (0,-0.8) {};
\draw[tran] (while) to node[font=\scriptsize,draw, fill=white, 
rectangle,pos=0.5] {$y<0$} (fin);
\draw[tran, loop, looseness = 5, in =-65, out = -115] (fin) to (fin);
\draw[tran] (while) -- node[font=\scriptsize,draw, fill=white, 
rectangle,pos=0.5, inner sep = 2pt] {$y\geq0$} node[label={[label 
distance=0.12cm]-90:{x:=...}}] {} (p1);
\draw[tran] (p1) to node[label={[label 
distance=0.12cm]-90:{y:=...}}] {} (p2);
\draw[tran] (lp) to  (p2);
\draw[tran] (p2) -- (p2|-dum2) -- node[font=\scriptsize,draw, fill=white, 
rectangle,pos=0.5, inner sep = 2pt] {$x\leq 0$} (lp|-dum2) -- (lp);
\draw[tran] (p2) -- (p2|-dum1) -- node[font=\scriptsize,draw, fill=white, 
rectangle,pos=0.5, inner sep = 2pt] {$x> 0$} (while|-dum1) -- (while);
%\draw[tran] (prob) to node[font=\scriptsize,draw, fill=white, 
%rectangle,pos=0.3, inner sep = 1pt] {$\frac{4}{10}$} (demon);
%\draw[tran] (demon) -- node[auto] {x:=x+1} (demon-|while) -- (while);
%\draw[tran] (angel) -- node[auto, swap] {x:=x+1} (angel-|while) -- (while);
%\draw[tran] (demon) to node[auto] {x:=x-1} (while);
%\draw[tran] (angel) to node[auto, swap] {x:=x-1} (while);

%\draw[tran] (angel) -- (aif);
%\draw[tran] (angel) -- (aelse);
%\draw[tran] (demon) -- (dif);
%\draw[tran] (demon) -- (delse);

%\draw[tran] (prob) -- node[font=\scriptsize,draw, fill=white, 
%rectangle,pos=0.5, inner sep = 1pt] {$\frac{3}{4}$} (prob|-dum1) -- node[auto] 
%{x:=x-1}
%(while|-dum1)--(while);
%\draw[tran] (prob) -- node[font=\scriptsize,draw, fill=white, 
%rectangle,pos=0.5, inner sep = 1pt] {$\frac{1}{4}$} (prob|-dum2) 
%-- node[auto,swap] {x:=x+1} (while|-dum2)--(while);
%\draw[tran] (dif) --  (demon-|while) -- node[auto, pos=0.2] {x:=x+1} (while);
%\draw[tran] (aif) --  (angel-|while) -- node[auto, swap, pos=0.2] {x:=x+1} 
%(while);
%\draw[tran] (delse) -- node[auto] {x:=x-1} (delse-|while.305) --  (while.-55);
%\draw[tran] (aelse) -- node[auto, swap] {x:=x-1} (aelse-|while.55) --  
%(while.55);
\end{tikzpicture}
\caption{A program with infinitely many reachable configurations which 
terminates with high probability, but not almost 
surely, together with a sketch of its pCFG. }\label{fig:nonterm}
\end{figure}

%We now show why this approach does not work for the probabilistic reachability and termination.

\smallskip
\begin{example}\label{ex:nonterm}
Consider the program in Figure~\ref{fig:nonterm}. 
In each 
iteration of the outer loop each of the variables is randomly modified by 
adding a number drawn from some uniform distribution.
%incremented or 
%decremented. The probability of incrementing $x$ is $\frac{3}{8}$, for
%decrementing $x$ it is $\frac{1}{8}$; the probability of incrementing $y$ it 
%is 
%$\frac{1}{8}$ and for decrementing $y$ it is $\frac{3}{8}$. 
Average increase of $x$ in each iteration is $\frac{3}{8}$, while average decrease 
of $y$ is $-\frac{3}{8}$.  It is easy to see that a program does not 
terminate almost-surely: 
there is for instance a tiny but non-zero probability of $x$ being decremented 
by at least $\frac{1}{8}$ in 
each of the first 240 loop iterations, after which we are stuck in the infinite 
inner 
loop. On 
the other hand, the expectations above show that there is a ``trend'' of $y$ 
decreasing and $x$ increasing, and executions that follow this trend eventually 
decrement $y$ below $0$ without entering the inner loop. Hence, the 
probabilistic intuition tells us that the program terminates with a high 
probability. However, the techniques of this section cannot prove 
this high-probability termination, since existence of an $\eps$-LRSM (with 
$\eps>0$) supported by a pure invariant already implies a.s. termination,
and so no such $\eps$-LRSM can exist for the program. 
\end{example}

In the next section we generalize the notion of pure invariants to stochastic invariants
for probabilistic termination to resolve issues like Example~\ref{ex:nonterm}.

%Adirect consequence of Theorems~\ref{thm:pure-supermart} is the following 
%connection 
%between pure invariants, 
%linear ranking supermartingales, and reachability properties. 
%
%\begin{corollary}
%\label{col:linear-ranking-old}
%Let $\lem$ be a linear $\eps$-ranking supermartingale for some set $C$ of 
%configurations of an \APP{} $P$ such that $\lem$ is supported by some pure 
%invariant of $P$. Then under every scheduler $\sigma$ it holds:
%
%\begin{compactenum}
%\item
%$\probm^{\sigma}(\treach{C}<\infty)=1$, i.e. the set $C$ is reached 
%almost-surely
%\item
%$\E^{\sigma}[\treach{C}]<\lem(\locinit,\vecinit)/\eps$.
%\item
%If $\{\lem(\cfg{\sigma}{i})\}_{i=0}^{\infty}$  has $c$-bounded differences for 
%some $c$, then 
%there exists a concentration bound $B$ for $\treach{C}$ such that $B\leq 
%({\lem(\locinit,\vecinit)}/{\eps})+1$.
%\end{compactenum}
%\end{corollary}

%\begin{proof}
%Denote $X_i = \lem(\cfg{\sigma}{i})$. To prove that 
%\end{proof}

%\begin{example}
%If we apply Theorem~\ref{thm:pure-supermart} on the $\frac{1}{4}$-LRSM $\lem$ 
%in Example~\ref{ex:lrsm}, we get exactly the process $\{X'_i\}_{i=0}^{\infty}$ 
%described in Example~\ref{ex:rsm}.
%\end{example}

%Having a supermartingale which decreases, on average, by some constant until 
%$C$ is reached effectively proves that $C$ is reached almost surely, very much like 
%the standard ranking functions prove that non-probabilistic program terminates. 
%Moreover, we can use such a supermartingale to prove additional assertions 
%about the reachability time.

%\begin{theorem}[variant of {\cite[Theorem 1]{CFNH16:prob-termination}}]
%\label{thm:old-ranking}
%Let $P$ be an \APP{}, $\sigma$ a scheduler, and 
%$(\Omega,\natfilt,\probm^\sigma)$ the corresponding probability space. Further, 
%let $C$ be a set of 
%configurations of $\pCFG_P$ such 
%that there exist an $\eps>0$ and an $\eps$-ranking supermartingale  
%$\{X_i\}_{i=0}^{\infty}$
%for $\treach{C}$. Then 
%\begin{compactenum}
%\item
%$\probm^{\sigma}(\treach{C}<\infty)=1$, i.e. the set $C$ is reached 
%almost-surely
%\item
%$\E^{\sigma}[\treach{C}]<\E^{\sigma}[X_0]/\eps$.
%\item
%If $\{X_i\}_{i=0}^{\infty}$  has $c$-bounded differences for some $c$, then 
%there exists a concentration bound $B$ for $\treach{C}$ such that $B\leq 
%({E^{\sigma}[X_0]}/{\eps})+1$.
%\end{compactenum}
%\end{theorem}

%\begin{remark}
%We again somewhat generalize the results of~\cite{CFNH16:prob-termination}, 
%where the 
%supermartingale 
%$\{X_i\}_{i=0}^{\infty}$ was additionally required to satisfy the 
%following: there exists $K>0$ such that with probability $1$ it holds $X_i > 
%-K$ for all $i$. We avoid this requirement by use of \emph{stopped ranking 
%supermartingale property} similar to~\cite[Section 10.9]{Williams:book}.
%\end{remark}

%Note that the requirement~\eqref{eq:ranking-sup} can be equivalently expressed 
%by stipulating that $\E[X_{i+1}|\mathcal{F}_i] \leq X_i - \eps\cdot 
%\vec{1}_{\{X_i \geq 0\}}$