### 2017-06-02 paper/lexicographic.tex
added smthing to the proof of LexRSM soundness

\section{Lexicographic Supermartingales}
\label{sec:lexicographic}

In this section we introduce the notion of a \emph{lexicographic ranking 
supermartingale}, which generalizes the standard notion of ranking 
supermartingales.

\begin{definition}[Lexicographic Ranking Supermartingale]
Let $(\Omega,\genfilt,\probm)$ be a probability space, 
$\{\genfilt_i\}_{i=0}^{\infty}$ a filtration of $\genfilt$, $\stime$ a stopping 
time w.r.t. that filtration, and 
$\eps\geq 0$. 
An $n$-dimensional stochastic process $\{\vec{X}_{i}\}_{i=0}^{\infty}$ is a 
\emph{lexicographic $\eps$-ranking supermartingale for $\stime$ 
($\eps$-LexRSM)} if the 
following 
conditions hold:
\begin{compactenum}
\item For each $1\leq j \leq n$ the 1-dimensional stochastic process 
$\{\vecseq{X}{i}{j}\}_{i=0}^{\infty}$ is adapted to 
$\{\genfilt_i\}_{i=0}^{\infty}$.
\item For each $i\in \Nset_0$ and each $1\leq j \leq n$ it holds 
$\E[|\vecseq{X}{i}{j}|]<\infty$.
\item For each $i\in \Nset_0$ and each $\omega\in \Omega$ with 
$\stime(\omega)>i$ there exists $1\leq j 
\leq n$ such that
\begin{compactitem}
	\item $\E[\vecseq{X}{i+1}{j}\mid 
	\genfilt_i](\omega)\leq\vecseq{X}{i}{j}(\omega) - 
	\eps$; and
	\item for all $1 \leq j' < j$ it holds $\E[\vecseq{X}{i+1}{j'}\mid 
	\genfilt_i](\omega)\leq\vecseq{X}{i}{j'}(\omega)$.
\end{compactitem}
\item For each $\omega \in \Omega$ and $i\in \Nset_0$ it holds 
$\stime(\omega)>i 
\Rightarrow \vec{X}_i (\omega)\geq \vec{0}$ (componentwise). Moreover, there 
exists a constant $K<0$ such that $\stime(\omega)=i 
\Rightarrow \vec{X}_i (\omega)\geq (K,\dots,K)$.
\end{compactenum}
\end{definition}

\begin{theorem}
Let $(\Omega,\genfilt,\probm)$ be a probability space, 
$\{\genfilt_i\}_{i=0}^{\infty}$ a filtration of $\genfilt$, and $\stime$ a 
stopping 
time w.r.t. that filtration. If there exists an $\eps$-LexRSM for $\stime$, 
then $\probm(\stime<\infty)=1$.
\end{theorem}
\begin{proof}
The proof proceeds by contradiction, i.e. we assume that an $\eps$-LexRSM for 
$\stime$ exists and $\probm(\stime=\infty)>0$. For succinctness we denote the 
set $\{\omega\mid \stime(\omega)=\infty\}$ by $\genRunSet_{\infty}$.

For $\omega\in \Omega$ we define the level of $\omega$ at step $i$ to be the 
number $\levelrank{\omega}{i}$ equal to either $0$, if $\stime(\omega)\leq i$, 
or otherwise to the smallest $1\leq j \leq n$ such that 
$\E[\vecseq{X}{i+1}{j}\mid 
\genfilt_i](\omega)\leq\vecseq{X}{i}{j}(\omega)-\eps$. By the definition of 
$\eps$-LexRSM the value $\levelrank{\omega}{i}$ is well-defined for all 
$\omega$. We denote by $\minlev(\omega)$ the smallest $0 \leq j \leq n$ such 
that $j$ is a level of $\omega$ at infinitely many steps. Note that $\omega \in 
\genRunSet_{\infty}$ if and only if $\minlev(\omega)\neq 0$. We denote by $M_i$ 
the set of all $\omega$'s with $\minlev(\omega)=i$.

Throughout the proof we use several times the following fundamental fact: if 
$\probm(A)>0$ for some set $A$ and $A=A_1\cup A_2 \cup A_3\cdots$ for some 
sequence of sets $A_1,A_2,A_3,\dots$, then there exists $i$ such that 
$\probm(A_i)>0$.

Now since $\genRunSet_{\infty}=M_1,\dots,M_n$, there must be $1\leq j \leq n$ 
s.t. $\probm(M_j)>0$. For each $\omega\in M_j$ there is the smallest number
$i_{\omega,j}\in\Nset_0$ such that for all $i\geq i_{\omega,j}$ it holds 
$\levelrank{\omega}{i}\geq 
j$. Denote by $S_{j,i}$ the set of all $\omega$'s in $M_j$ s.t. 
$i_{\omega,j}=i$. 
Since $M_j = M_{j,1} \cup M_{j,2} \cup M_{j,3} \cup \cdots$, there is $i\in 
\Nset_0$ 
s.t. 
$\probm(M_{j,i})>0$. Continuing on the same note, for each $B\in \Nset$ we 
denote by $M_{j,i}^{B}$ the set off all $\omega$'s in $M_{j,i}$ s.t. 
$\vecseq{X}{i}{j}(\omega)\leq B$. Since $M_{j,i}= M_{j,i}^{1} \cup M_{j,i}^{2} 
\cup M_{j,i}^{3} \cup \cdots  $, there is $B\in \Nset$ s.t. 
$\probm(M_{j,i}^{B})>0$.

%\begin{fact}

%\end{fact}

%\[
%\levelrank{\omega}{i} = \begin{dcases}
%0 & \text{if $\stime(\omega)\leq i$} \\
% 
%\end{dcases}
%\]
\end{proof}