### 2017-06-11 paper/lexicographic.tex

\section{Lexicographic Supermartingales}
\label{sec:lexicographic}

In this section we introduce the notion of a \emph{lexicographic ranking 
supermartingale}, which generalizes the standard notion of ranking 
supermartingales.

\begin{definition}[Lexicographic Ranking Supermartingale]
Let $(\Omega,\genfilt,\probm)$ be a probability space, 
$\{\genfilt_i\}_{i=0}^{\infty}$ a filtration of $\genfilt$, $\stime$ a stopping 
time w.r.t. that filtration, and 
$\eps\geq 0$. 
An $n$-dimensional stochastic process $\{\vec{X}_{i}\}_{i=0}^{\infty}$ is a 
\emph{lexicographic $\eps$-ranking supermartingale for $\stime$ 
($\eps$-LexRSM)} if the 
following 
conditions hold:
\begin{compactenum}
\item For each $1\leq j \leq n$ the 1-dimensional stochastic process 
$\{\vecseq{X}{i}{j}\}_{i=0}^{\infty}$ is adapted to 
$\{\genfilt_i\}_{i=0}^{\infty}$.
%\item For each $i\in \Nset_0$ and each $1\leq j \leq n$ it holds 
%$\E[|\vecseq{X}{i}{j}|]<\infty$.
\item For each $i\in \Nset_0$ and each $\omega\in \Omega$ with 
$\stime(\omega)>i$ there exists $1\leq j 
\leq n$ such that
\begin{compactitem}
	\item $\E[\vecseq{X}{i+1}{j}\mid 
	\genfilt_i](\omega)\leq\vecseq{X}{i}{j}(\omega) - 
	\eps$; and
	\item for all $1 \leq j' < j$ it holds $\E[\vecseq{X}{i+1}{j'}\mid 
	\genfilt_i](\omega)\leq\vecseq{X}{i}{j'}(\omega)$.
\end{compactitem}
\item For each $\omega \in \Omega$ and $i\in \Nset_0$ it holds 
$\vec{X}_i (\omega)\geq \vec{0}$ (componentwise).
\end{compactenum}
\end{definition}

\begin{theorem}
Let $(\Omega,\genfilt,\probm)$ be a probability space, 
$\{\genfilt_i\}_{i=0}^{\infty}$ a filtration of $\genfilt$, and $\stime$ a 
stopping 
time w.r.t. that filtration. If there exists an $\eps$-LexRSM for $\stime$, 
then $\probm(\stime<\infty)=1$.
\end{theorem}
\begin{proof}
The proof proceeds by contradiction, i.e. we assume that an $\eps$-LexRSM for 
$\stime$ exists and $\probm(\stime=\infty)>0$. For succinctness we denote the 
set $\{\omega\mid \stime(\omega)=\infty\}$ by $\genRunSet_{\infty}$.

For $\omega\in \Omega$ we define the level of $\omega$ at step $i$ to be the 
number $\levelrank{\omega}{i}$ equal to either $0$, if $\stime(\omega)\leq i$, 
or otherwise to the smallest $1\leq j \leq n$ such that 
$\E[\vecseq{X}{i+1}{j}\mid 
\genfilt_i](\omega)\leq\vecseq{X}{i}{j}(\omega)-\eps$. By the definition of 
$\eps$-LexRSM the value $\levelrank{\omega}{i}$ is well-defined for all 
$\omega$ and moreover, the random variable $\levelrank{}{i}$ is $\genfilt_i$-measurable. We denote by $\minlev(\omega)$ the smallest $0 \leq j \leq n$ such 
that $j$ is a level of $\omega$ at infinitely many steps. Note that $\omega \in 
\genRunSet_{\infty}$ if and only if $\minlev(\omega)\neq 0$. We denote by $M_i$ 
the set of all $\omega$'s with $\minlev(\omega)=i$.

Throughout the proof we use several times the following fundamental fact: if 
$\probm(A)>0$ for some set $A$ and $A=A_1\cup A_2 \cup A_3\cdots$ for some 
sequence of sets $A_1,A_2,A_3,\dots$, then there exists $i$ such that 
$\probm(A_i)>0$.

Now since $\genRunSet_{\infty}=M_1,\dots,M_n$, there must be $1\leq \fixn{j} \leq n$ 
s.t. $\probm(M_{\fixn{j}})>0$. For each $\omega\in M_{\fixn{j}}$ there is the smallest number
${i}_{\omega,\fixn{j}}\in\Nset_0$ such that for all $i\geq {i}_{\omega,{j}}$ it holds 
$\levelrank{\omega}{i}\geq \fixn{j}$. Denote by $S_{\fixn{j},{i}}$ the set of all $\omega$'s in $M_{\fixn{j}}$ s.t. 
$i_{\omega,\fixn{j}}={i}$. 
Since $M_{\fixn{j}} = M_{\fixn{j},1} \cup M_{\fixn{j},2} \cup M_{\fixn{j},3} \cup \cdots$, there is $\fixn{i}\in 
\Nset_0$ 
s.t. 
$\probm(M_{\fixn{j},\fixn{i}})>0$. Continuing on the same note, for each $B\in \Nset$ we 
denote by $M_{\fixn{j},\fixn{i}}^{B}$ the set off all $\omega$'s in $M_{\fixn{j},\fixn{i}}$ s.t. 
$\vecseq{X}{\fixn{i}}{\fixn{j}}(\omega)\leq B$. Since $M_{\fixn{j},\fixn{i}}= M_{\fixn{j},\fixn{i}}^{1} \cup M_{\fixn{j},\fixn{i}}^{2} 
\cup M_{\fixn{j},\fixn{i}}^{3} \cup \cdots  $, there is $\fixn{B}\in \Nset$ s.t. 
$\probm(M_{\fixn{j},\fixn{i}}^{\fixn{B}})>0$. Denote $\fixn{p}=\probm(M_{\fixn{j},\fixn{i}}^{\fixn{B}})$.

Let $D$ be the set of all $\omega\in \Omega$ such that 
$\vecseq{X}{\fixn{i}}{\fixn{j}}(\omega)\leq \fixn{B}$. Note that $M_{\fixn{j},\fixn{i}}^{B} \subseteq D$ and $D\in 
\genfilt_i$ (and thus also $D\in \genfilt_{i'}$ for all $i'\geq i$). 
Define a stopping time $U$ w.r.t. filtration $\{\genfilt_i\}_{i=0}^{\infty}$ as follows: for all $\omega \in \Omega$ 
we put 
$U(\omega)=\inf\{k\in \Nset_0 \mid k\geq \fixn{i} \text{ and }
\levelrank{\omega}{k} < \fixn{j} \}$.

Define a 
(one-dimensional) stochastic process $\{Y_k\}_{k=0}^{\infty}$ as 
follows:
$$
Y_k(\omega) = \begin{dcases}
0 & \text{if $\omega \not\in D$}\\
\fixn{B} & \text{if $\omega \in D$ and $k<\fixn{i}$ }\\
\vecseq{X}{k}{\fixn{j}}(\omega) & \text{if $\omega \in D$, $k\geq \fixn{i}$ and 
$U(\omega)>k$ 
}\\
\vecseq{X}{U(\omega)}{\fixn{j}}(\omega) & \text{if $\omega\in D$, $k\geq 
\fixn{i}$ and 
$U(\omega)\leq k$ }.
\end{dcases}
$$

We prove several properties of the process $\{Y_k\}_{k=0}^{\infty}$. 
First, clearly for all $k\geq 0$, $Y_k(\omega)\geq 0$. Second, for each $k\geq \fixn{i}$, the variable $Y_k$ is $\genfilt_k$-measurable, as $D\in \genfilt_{\fixn{i}}$, $\{\vecseq{X}{i}{\fixn{j}}(\omega)\}_{i=0}^{\infty}$ is adapted to the filtration $\{\genfilt_i\}_{i=0}^{\infty}$ and $U$ is a stopping time w.r.t. this filtration. Finally, for any $k\in \Nset_0$ denote by 
$\noofdec_k$ the random variable such that $\noofdec_k(\omega) = |\{i'\in 
\Nset\mid \fixn{i} \leq i' < k \text{ and } \levelrank{\omega}{i'}=\fixn{j}\} |$.
We claim 
that for each $k\geq \fixn{i}$ 
it holds 
\begin{equation}
\label{eq:lexrsm-soundness-main}
\E[Y_k]\leq \fixn{B}\cdot \probm(D) - \eps\cdot\sum_{\ell=0}^{k-\fixn{i}} \ell\cdot\probm(D 
\cap \{U\geq k\} \cap \{\noofdec_k 
= 
\ell\}).
\end{equation}

The proof of~\eqref{eq:lexrsm-soundness-main} goes by induction on $k$. For 
$k=\fixn{i}$ the sum on the right-hand side equals $0$, so the 
inequality immediately follows from the definition of $Y_k$. Now assume 
that~\eqref{eq:lexrsm-soundness-main} holds for some $k\geq \fixn{i}$. We have that 
\begin{align}
\label{eq:lexrsm-ind-1}
\E[Y_{k+1}] &= \underbrace{\E[Y_{k+1}\cdot \indicator{\Omega\setminus D}]}_{=0=\E[Y_k\cdot\indicator{\Omega\setminus D}]} +  \underbrace{\E[Y_{k+1}\cdot \indicator{D \cap \{U \leq k\}}]}_{=\E[Y_k\cdot\indicator{D\cap \{U\leq k\}}]} + \underbrace{\E[Y_{k+1}\cdot \indicator{D \cap \{U > k\}}]}_{=\E[\vecseq{X}{k+1}{\fixn{j}}\cdot\indicator{D\cap \{U>k \} }]},
\end{align}
where the equality $\E[Y_{k+1}\cdot \indicator{D \cap \{U \leq k\}}] =\E[Y_k\cdot\indicator{D\cap \{U\leq k\}}] $ follows from the fact that $Y_{k+1}(\omega)=Y_k({\omega})=\vecseq{X}{U(\omega)}{\fixn{j}}(\omega)$ for $\omega\in \{U\leq k\}$, and similarly for the last term. We prove that 
\begin{equation}
\label{eq:lexrsm-ind-2}
\E[\vecseq{X}{k+1}{\fixn{j}}\cdot \indicator{D\cap \{U > k \}}] \leq \E[Y_k\cdot \indicator{D\cap \{U > k \}} -\eps\cdot \indicator{D\cap \{U>k\} \cap \{\levelrank{}{k}= \fixn{j}\} }] .
\end{equation}
Indeed, it holds
\begin{equation}
\label{eq:lexrsm-ind-3}
\E[\vecseq{X}{k+1}{\fixn{j}}\cdot \indicator{D\cap \{U > k \}}] = \E[\vecseq{X}{k+1}{\fixn{j}}\cdot \indicator{D\cap \{U > k \} \cap \{\levelrank{}{k}=\fixn{j} \} }]  + \E[\vecseq{X}{k+1}{\fixn{j}}\cdot \indicator{D\cap \{U > k \} \cap \{\levelrank{}{k}>\fixn{j}\} }], 
\end{equation}
since $\levelrank{\omega}{k}\geq \fixn{j}$ for all $\omega \in \{U>k\}$. Since the set $D\cap \{U > k \} \cap \{\levelrank{}{k}=\fixn{j} \}$ is $\genfilt_k$-measurable, we get
\begin{align}
\E[\vecseq{X}{k+1}{\fixn{j}}\cdot \indicator{D\cap \{U > k \} \cap \{\levelrank{}{k}=\fixn{j} \} }] &= \E[\E[\vecseq{X}{k+1}{\fixn{j}}\mid \genfilt_k]\cdot \indicator{D\cap \{U > k \} \cap \{\levelrank{}{k}=\fixn{j} \} }] \label{eq:lexrsm-ind-4}\\
&\leq \E[(\vecseq{X}{k}{\fixn{j}} - \eps)\cdot \indicator{D\cap \{U > k \} \cap \{\levelrank{}{k}=\fixn{j} \}}] \label{eq:lexrsm-ind-5}\\
&=\E[(Y_k - \eps)\cdot \indicator{D\cap \{U > k \} \cap \{\levelrank{}{k}=\fixn{j} \}}] \label{eq:lexrsm-ind-6},
\end{align}
where~\eqref{eq:lexrsm-ind-4} follows from the definition of conditional expectation~\eqref{eq:cond-exp},~\eqref{eq:lexrsm-ind-5} follows from the definition of $\{\levelrank{}{k}=\fixn{j}\}$, and~\eqref{eq:lexrsm-ind-6} holds since $Y_k(\omega)=\vecseq{X}{k}{\fixn{j}}(\omega)$ for $\omega$ with $U(\omega)> k$. Almost identical argument shows that
\begin{equation}
\label{eq:lexrsm-ind-7}
\E[\vecseq{X}{k+1}{\fixn{j}}\cdot \indicator{D\cap \{U > k \} \cap \{\levelrank{}{k}>\fixn{j} \} }] \leq \E[Y_k\cdot \indicator{D\cap \{U > k \} \cap \{\levelrank{}{k}>\fixn{j} \}}].
\end{equation}
Plugging~\eqref{eq:lexrsm-ind-6} and~\eqref{eq:lexrsm-ind-7} into~\eqref{eq:lexrsm-ind-3} yields~\eqref{eq:lexrsm-ind-2}. Now we can plug~\eqref{eq:lexrsm-ind-2} into~\eqref{eq:lexrsm-ind-1} to get
\begin{align}
\E[Y_{k+1}]&\leq \E[Y_k] - \eps\cdot \E[\indicator{D \cap \{U>k\} \cap \{\levelrank{}{k} = \fixn{j} \} }] = \E[Y_k] - \eps\cdot \probm(D \cap \{U>k\} \cap \{\levelrank{}{k} = \fixn{j} \} ) \nonumber\\
&\leq \fixn{B}\cdot \probm(D) - \eps\cdot\left(\sum_{\ell=0}^{k-\fixn{i}} \ell\cdot\probm(D 
\cap \{U\geq k\} \cap \{\noofdec_k = \ell\})\right) -\eps\cdot \probm(D \cap \{U>k\} \cap \{\levelrank{}{k} = \fixn{j} \}),
\end{align}
where the last inequality follows from induction hypothesis. Hence, using $D_{k,\ell}$ as a shorthand for $D 
\cap \{U\geq k\} \cap \{\noofdec_k = \ell\}$, to prove~\eqref{eq:lexrsm-soundness-main} it remains to show that
\begin{equation}
\label{eq:lexrsm-ind-8}
\sum_{\ell=0}^{k-\fixn{i}} \ell\cdot\probm(D_{k,\ell}) + \probm(D \cap \{U>k\} \cap \{\levelrank{}{k} = \fixn{j} \}) = \sum_{\ell=0}^{k+1-\fixn{i}} \ell\cdot\probm(D_{k+1,\ell}).
\end{equation}
The left-hand side of~\eqref{eq:lexrsm-ind-8} is equal to
\begin{align}
&\phantom{+}\;\sum_{\ell=0}^{k-\fixn{i}} \ell\cdot\probm(D_{k,\ell} \cap \{\levelrank{}{k}=\fixn{j}\}) + \sum_{\ell=0}^{k-\fixn{i}} \ell\cdot\probm(D_{k,\ell}\cap \{\levelrank{}{k}>\fixn{j}\} )\nonumber \\ 
&+\sum_{\ell=0}^{k-\fixn{i}}\probm(\underbrace{D \cap \{U>k\} \cap \{\levelrank{}{k} = \fixn{j}\} \cap \{\noofdec_k = \ell\}}_{=D_{k,\ell} \cap \{\levelrank{}{k} = \fixn{j}\}})
\nonumber \\
&= \sum_{\ell=0}^{k-\fixn{i}}(\ell+1) \cdot\probm(D_{k,\ell}\cap \{\levelrank{}{k} = \fixn{j}\}) +  \sum_{\ell=0}^{k-\fixn{i}} \ell\cdot\probm(D_{k,\ell}\cap \{\levelrank{}{k}>\fixn{j}\} ) \nonumber\\
&=\sum_{\ell=0}^{k-\fixn{i}} (\ell+1)\cdot \probm{(D_{k+1,\ell+1} \cap \{\levelrank{}{k} = \fixn{j}\})} +  \sum_{\ell=0}^{k-\fixn{i}} \ell\cdot\probm(D_{k+1,\ell}\cap \{\levelrank{}{k}>\fixn{j}\} ) \label{eq:lexrsm-ind-9}\\
&= \sum_{\ell=1}^{k+1-\fixn{i}} \ell\cdot \probm{(D_{k+1,\ell} \cap \{\levelrank{}{k} = \fixn{j}\})} +  \sum_{\ell=0}^{k-\fixn{i}} \ell\cdot\probm(D_{k+1,\ell}\cap \{\levelrank{}{k}>\fixn{j}\} )
   \nonumber \\
&= (k+1-\fixn{i})\cdot \probm{(D_{k+1,k+1-\fixn{i}}\cap\{\levelrank{}{k} =\fixn{j} \} )} + \sum_{\ell=1}^{k-\fixn{i}} \ell \cdot \probm(D_{k+1,\ell})  \nonumber\\
&= (k+1-\fixn{i})\cdot\probm(D_{k+1,k+1-\fixn{i}}) + \sum_{\ell=1}^{k-\fixn{i}}\ell\cdot\probm(D_{k+1,\ell}) = \text{ right-hand side of~\eqref{eq:lexrsm-ind-8}} \label{eq:lexrsm-ind-11}.
\end{align}
The individual steps in the above computation are justified as follows: in~\eqref{eq:lexrsm-ind-9} we use the facts that for all $\omega$'s whose level in step $k$ is $\fixn{j}$ it holds that $\noofdec_k(\omega)+1=\noofdec_{k+1}(\omega)$, and similarly, for $\omega$'s whose level in step $k$ is $>\fixn{j}$ it holds $\noofdec_k(\omega)=\noofdec_{k+1}(\omega)$. Moreover, for all $\omega\in D_{k,\ell}$ it holds that $\levelrank{k}{\omega}\geq \fixn{j} \Rightarrow U(\omega)\geq k+1$. Finally, in~\eqref{eq:lexrsm-ind-11} we use the fact that all $\omega\in D_{k+1,k+1-\fixn{i}}$ need to have level $\fixn{j}$ in step $k$, since otherwise such an $\omega$ would need to have level $\fixn{j}$ for at least $k+1-\fixn{i}$ times within steps $\{\fixn{i},\fixn{i}+1,\dots,k-1\}$, but there are $k-\fixn{i}$ such steps, a contradiction. This concludes the proof of~\eqref{eq:lexrsm-soundness-main}.

Now according to~\eqref{eq:lexrsm-soundness-main} it holds  $\E[Y_k]\leq \fixn{B}\cdot \probm(D) - \eps\cdot\sum_{\ell=0}^{k-\fixn{i}} \ell\cdot\probm(D 
\cap \{U\geq k\} \cap \{\noofdec_k 
= 
\ell\})$ for all $k\geq \fixn{i}$. Let $m=3\fixn{B}\cdot \probm(D)/(\eps\cdot \fixn{p})$. For each $\omega\in M_{\fixn{j},\fixn{i}}^{\fixn{B}} $ there exists $k_m(\omega)\geq \fixn{i}$ such that $\noofdec_{k_{m}(\omega)}\geq m$. Hence, $M_{\fixn{j},\fixn{i}}^{\fixn{B}} = \bigcup_{k=\fixn{i}}^{\infty} M_{\fixn{j},\fixn{i}}^{\fixn{B}} \cap \{k_{m}(\omega) \leq k\}$ and $\fixn{p}=\probm(M_{\fixn{j},\fixn{i}}^{\fixn{B}})=\lim_{k\rightarrow \infty}\probm(M_{\fixn{j},\fixn{i}}^{\fixn{B}} \cap \{k_{m} \leq k\})$. Thus, there exists $k_0$ such that $\probm(M_{\fixn{j},\fixn{i}}^{\fixn{B}} \cap \{\noofdec_{k_0}\geq m\}) \geq \fixn{p}/2$. Denote $\fixn{M}= M_{\fixn{j},\fixn{i}}^{\fixn{B}} \cap \{\noofdec_{k_0}\geq m\}$. Clearly $\fixn{M} \subseteq D\cap \{U\geq k_0 \} \cap \{\noofdec_{k_0}\geq m \}$. From \eqref{eq:lexrsm-soundness-main} it follows that
\begin{align*}
\E[Y_{k_0}]&\leq \fixn{B}\cdot \probm(D) - \eps\cdot\sum_{\ell=0}^{k_0-\fixn{i}} \ell\cdot\probm(D 
\cap \{U\geq k_0\} \cap \{\noofdec_{k_0} 
=
\ell\}) \\
&= \fixn{B}\cdot \probm(D) - \eps\cdot \sum_{\ell=1}^{k_0-\fixn{i}} \probm(D 
\cap \{U\geq k_0\} \cap \{\noofdec_{k_0} 
\geq 
\ell\})\\
& \leq \fixn{B}\cdot \probm(D) - \eps\cdot \sum_{\ell=1}^{m} \probm(D 
\cap \{U\geq k_0\} \cap \{\noofdec_{k_0} 
\geq 
\ell\})\\
&\leq \fixn{B}\cdot \probm(D) - \eps\cdot m \cdot \fixn{p}/2 < 0, 
\end{align*}
the last inequality following by substituting for $m$. But for each $k$ the random variable $Y_k$, so it must also have a non-negative expectation, a contradiction.
%\begin{enumerate}
%\
%\end{itemize}
%\end{enumerate}

%\begin{fact}

%\end{fact}

%\[
%\levelrank{\omega}{i} = \begin{dcases}
%0 & \text{if $\stime(\omega)\leq i$} \\
% 
%\end{dcases}
%\]
\end{proof}