### 2017-06-23 paper/compositional.tex

\section{Compositionality of Ranking Supermartingales Revisited}
\label{sec:compositional}

Compositionality in the context of termination proving means providing the 
proof of termination step-by-step, handling one loop at a time, rather than 
attempting to construct the proof (in our case, a LexRSM) at once, by solving a 
single large system of constraints. In non-probabilistic setting, the 
relationship between lexicographic ranking functions and compositional 
termination proving is well established, as well as the importance of 
compositional proofs for efficient implementation~\cite{xxx}. In the context of 
probabilistic programs, the work~\cite{HolgerPOPL} attempted to provide a 
compositional notion of almost-sure termination proof based on the 
\emph{probabilistic variant rule (V-rule),} which we explain in a more detail 
below. 
However, as already noted in~\cite{HolgerPOPL}, the probabilistic V-rule is not 
sound in general. In order to rectify this issue, in~\cite{HolgerPOPL} they 
impose an additional constraint of \emph{uniform integrability} on ranking 
supermartingales that prove termination of individual loops, 
under which the probabilistic V-rule is shown to be sound. Apart from uniform 
integrability being somewhat restrictive in itself, in~\cite{HolgerPOPL} it is 
argued that proving uniform integrability is beyond the capability of 
state-of-the-art automated theorem provers. As a substitute for 
these,~\cite{HolgerPOPL} introduces a type system that can be used to 
automatically prove 
uniform integrability of ranking supermartingales for a restricted class of 
programs. In particular, the method cannot handle programs in which termination 
is controlled by variable that can be modified by a non-constant value, e.g. 
variable $x$ which appears in an assignment of the form $x:=x-y$. In this 
section we show that using our insights in lexicographic supermartingales we 
can obtain a subtly different notion of a probabilistic V-rule which is sound 
without any additional assumptions, and which can be used to prove termination 
of programs that the previous method cannot handle.

Let $\program$ be a \PP{} of the form $\textbf{while } \Psi \textbf{ do } 
\programbody \textbf{ od}$, and let $\pCFG_{\program}$ be the associated pCFG, whose set of locations we denote by $\locs$. We denote by $\loops(\program)$ the set of all locations of $\pCFG_{\program}$ that belong to a sub-pCFG of $\pCFG_{\program}$ corresponding to some nested loop of $\program$. We also define $\slice(\program)$ to be the set $\locs\setminus\loops(\program)$ of locations that do not belong to any nested sub-loop.  A formal definition of 
both functions is given in \AppendixMaterial, we illustrate them in the 
following example.
%We denote by $\loops(\program)$ the set of all 
%loops in $\program$ there are on the top level of $\programbody$, i.e. they are 
%nested loops of $\program$ nested directly "below" the outer loop. We define an 
%un-nested slice of $\program$ to be a \PP{} $\slice(\program)$ obtained from 
%$\program$ by removing all loops in $\loops(\program)$ and replacing them with 
%skip statements. We also define the set 
%$\alloops(\program)$ of all loops in $\program$, i.e. 
%$\alloops(\program):=\loops(\program)\cup \{\loops(\program')\mid \program'\in 
%\loops(\program)\}$. The above definitions do not specify the initialization 
%preamble for the sub-programs $\program'\in\alloops(\program)$: we can deem 
%this pre-amble empty, i.e. consider these sub-programs with any initial 
%valuation, or, provided that some invariant map $\inv$ for $\program$ is 
%given, we can set the set of possible initial valuations of a sub-program 
%$\program'$ to be the set $\inv(\loc^{\lin}_{\program'})$, where 
%$\inv(\loc^{\lin}_{\program'})$ is the location of $\program$ corresponding to 
%the head of the loop $\program'$.
%A formal definition of 
%both functions is given in \AppendixMaterial, we illustrate them in the 
%following example.

\textbf{[PETR: INSERT EXAMPLE]}

We say that a 1-dimensional measurable map $\lem$ for $\pCFG_{\program}$ is $\eps$-ranking in location $\loc$, if for each $\vec{x}\in\Rset^{|\pvars|}$ such that $(\loc,\vec{x}) $ is reachable from some initial configuration it holds $\preexp{\lem}(\loc,\vec{x})\leq \lem(\loc,\vec{x})-\eps.$ The map $\lem$ is \emph{unaffected} in location $\loc$ if it is $0$-ranking in $\loc$, and non-negative in $\loc$ if $\lem(\loc,\vec{x})\geq 0$ for all $\vec{x}$ such that $(\loc,\vec{x})$ is reachable from an initial configuration. A measurable map $\lem$ is propositionally linear, if each function $\lem(\loc)$ is of the form $\indicator{G_1}\cdot E_1 + \cdots \indicator{G_k}\cdot E_k$, where each $\indicator{G_i}$ is an indicator function of some polyhedron and each $E_i$ is a linear expression.

We recall the notion of compositional ranking supermartingale as introduced in~\cite{HolgerPOPL}. We call it a PV supermartingale, as it is based on so called probabilistic variant rule. Due to differences in syntax and semantics, the definition is syntactically slightly different from~\cite{HolgerPOPL}, but the essence is the same.

\begin{definition}[PV-supermartingale]
A 1-dimensional propositionally linear map $\lem$ is a PV supermartingale (PVSM) for a program $\program$ if there exists $\eps>0$ such that $\lem$ is $\eps$-ranking an non-negative in each location $\loc\in\slice{(\program)}$ and unaffected in each $\loc\in\loops(\program)$.
\end{definition}

In~\cite{HolgerPOPL} they show that even if all nested loops were already proved to terminate a.s. and there is a PVSM for the program, then the program itself might not terminate a.s. Then they investigate conditions under which existence of a PVSM together with a proof of a.s. termination of each nested sub-loop of $\program$ entails termination of the whole program $\program$. The condition were formulated in terms of a certain type system for \PP{}s, so we say that a PVSM satisfying the condition \emph{typechecks} correctly. In~\cite{HolgerPOPL} the following was proved:

\begin{theorem}[\cite{HolgerPOPL}]
\label{thm:holger-comp}
Let $\program$ be a \PP{} of the form $\textbf{while } \Psi \textbf{ do } 
\programbody \textbf{ od}$. Assume that each nested loop of $\program$ terminates almost surely from each reachable configuration, and that there exists a PVSM for $\program$ that typechecks correctly. Then $\program$ terminates almost surely.
\end{theorem}

We do not repeat the precise definition of the typesystem here, we just explain its motivation and give examples where it fails, i.e. examples where $\program$ terminates a.s. (and hence the same holds also for its nested loops) and a PV martingale exists, but no PVSM typechecks.

The purpose of the typesystem is, in principle, to ensure that the stochastic process which, in each time step, returns the current value of the PVSM, is uniformly integrable inside the nested loops. Uniform integrability is a deep concept from probability and measure theory: a sequence $X_0,X_1,X_2,\dots$ of random variables is uniformly integrable if for each $\delta>0$ there exists an $K\in \Nset$ such that for all $n\geq 0$ it holds $\E[|X_n|\cdot\indicator{X_n \geq K}]\leq \delta$. As noted in~\cite{HolgerPOPL}, proofs of uniform integrability are hard to (semi)-automatize, and hence they devise the typechecking as a sound (but not complete) method of proving uniform integrability. Take for example \textbf{[PETR: ADD EXAMPLE]}. However, as we show below, for none of these programs there exists a PVSM that typechecks.

Indeed, taking a closer look at typesystem in~\cite{HolgerPOPL}, there are several reasons for typechecking of PVSM to fail. The major ones are:
\begin{compactenum}
\item
A PVSM $\lem$ for \PP{} $\program$ will not typecheck if $\program$ has a nested loop in which the value of $\lem$ can change unboundedly in a single step (see Example~\ref{xxx}). 
\item
A PVSM $\lem$ for  \PP{} $\program$ will not typecheck if $\program$ has a nested loop which itself has a nested loop in which some variable appearing in some expression in $\lem$ is modified, see Example~\ref{xxx}.\footnote{Both these statements regarding typechecking failure are somewhat simplified, even in these two cases the PVSM might sometimes typecheck correctly, in case where the nested loops are followed by assignments which completely overwrite the effect of these loops, e.g. if \textbf{[PETR: POINTER TO AN EXAMPLE]}. However, the statements intuitively summarize the major reasons for typechecking failure. }
\end{compactenum}

To overcome the rather strict

%Each measurable map for $\lem$ can be, in a natural way, restricted to a 
%measurable map $\lem_{\program'}$ for $\program'\in \loops(\program)$, 
%since each pCFG $\pCFG_{\program'}$ is a sub-pCFG of $\pCFG_{\program}$. 
%Similarly, a measurable map $\lem$ for $\program$ can be restricted to a 
%measurable map $\lem_{\slice}$ for $\slice(\program)$, since 
%$\pCFG_{\slice(\program)}$ is produced from $\pCFG_{\program}$ essentially by 
%removing certain locations (provided that we identify the locations forming 
%heads of nested loops with locations preceding skip statements), and 
%$\lem_{\slice}$ is defined to be equal to $\lem$ in all the locations that 
%remained. Similarly,