### 2017-07-07 paper/exp_appendix.tex
Changes in table

\section{Experimental Results}\label{sec:app_ex}
% The table which is to be there in appendix
In Table~\ref{tab:exp1} in Section~\ref{sec:experiments} we presented results for twenty five benchmarks.
We now present additional result for other twenty seven benchmarks in Table~\ref{tab:exp3}.

\smallskip\noindent{\em Additional description of table.}
For the benchmark results reported in Table~\ref{tab:exp1} and Table~\ref{tab:exp3}, 
we explain the possibilities of Yes/No in the last two column entries.
\begin{compactitem}
\item If for the non-probabilistic program no lexicographic RSM exists, then the last two 
column entries are no and no.

\item If extending the probabilistic program with both probabilistic loop and assignment,
a lexicographic RSM exists, then the last two column entries are yes and yes.

\item If there exist lexicographic RSM for the non-probabilistic program, but there is no
lexicographic RSM for the program extended with probabilistic loops, then the second-last 
entry is yes and the last entry is no, and no RSM exists (i.e., the third entry is no).

\item If there exists lexicographic RSM for extending the non-probabilistic program 
with probabilistic loop, but no lexicographic RSM for extending the non-probabilistic programs
with both probabilistic loops and assignments, then for such benchmarks there are two rows 
explaining the results. 
Hence while in all previous cases there is a single row for the benchmarks, for this case,
there are two rows presenting the results.
\end{compactitem}

\begin{center}
\begin{table}[]
  \centering
   \begin{tabular}{c|c|c|c|c|c}

{Benchmark} & {Time (s)} & {Solution} & {Dimension} & {Prob. loops} & {Prob. Assignments} \\\hline \hline
{aaron2} & {0.09} & {yes} & {2} & {yes} & {yes} \\\hline
{ax} & {0.11} & {yes} & {3} & {yes} & {yes} \\\hline
{complex} & {0.10} & {yes} & {1} & {yes} & {yes} \\\hline
{counterex1b} & {0.10} & {yes} & {3} & {yes} & {yes} \\\hline
{cousot9} & {0.09} & {no} & {-} & {yes} & {no} \\\hline
{easy2} & {0.09} & {yes} & {2} & {yes} & {yes} \\\hline
{loops} & {0.10} & {no} & {-} & {no} & {no} \\\hline
{\multirow{2}{*}{nestedloop}} & {0.16} & {yes} & {3} & {yes} & {no} \\\cline{2-6}
{} & {0.23} & {no} & {-} & {yes} & {yes} \\\hline
{perfect1} & {0.10} & {yes} & {3} & {yes} & {yes} \\\hline
{random1d} & {0.10} & {yes} & {2} & {yes} & {yes} \\\hline
%{realbubble} & {0.22} & {yes} & {3} & {yes} & {yes} \\\hline
{realheapsort} & {0.22} & {no} & {-} & {no} & {no} \\\hline
{realheapsort\_step1} & {0.11} & {no} & {-} & {no} & {no} \\\hline
{realheapsort\_step2} & {0.16} & {no} & {-} & {yes} & {no} \\\hline
{rsd} & {0.10} & {yes} & {1} & {yes} & {yes} \\\hline
{sipma91} & {0.14} & {yes} & {2} & {yes} & {yes} \\\hline
{speedFails1} & {0.09} & {yes} & {2} & {yes} & {yes} \\\hline
{speedDis1} & {0.09} & {no} & {-} & {no} & {no} \\\hline
{\multirow{2}{*}{speedFails2}} & {0.14} & {yes} & {1} & {yes} & {no} \\\cline{2-6}
{} & {0.08} & {no} & {-} & {yes} & {yes} \\\hline
{speedFails4} & {0.09} & {no} & {-} & {no} & {no} \\\hline
{speedNestedMultipleDep} & {0.10} & {yes} & {3} & {yes} & {yes} \\\hline
{speedpldi3} & {0.10} & {yes} & {3} & {yes} & {yes} \\\hline
{speedSimpleMultiple} & {0.10} & {no} & {-} & {no} & {no} \\\hline
{speedSingleSingle} & {0.11} & {yes} & {2} & {yes} & {yes} \\\hline
{terminate} & {0.09} & {yes} & {2} & {yes} & {yes} \\\hline
{wcet0} & {0.11} & {yes} & {2} & {yes} & {yes} \\\hline
{wcet2} & {0.09} & {yes} & {2} & {yes} & {yes} \\\hline
{wise} & {0.09} & {no} & {-} & {no} & {no} \\\hline
    \end{tabular}
    \caption{Additonal experimental results for benchmarks from~\cite{ADFG10:lexicographic} extended with probabilistic loops/probabilistic assignments.}
    \label{tab:exp3}
\end{table}
\end{center}

% The table for constructed long examples